---
title: "Computational Modeling - Assignment 2"
author: "Riccardo Fusaroli"
date: "29/01/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


```

## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci

N.B. there is a second part at the bottom for next week.

### First part

You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

To keep things simple (your teachers should not be faced with too complicated things):
- You created a pool of equally challenging questions on CogSci
- Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
- Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

This is the data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

Questions:

1. What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5) [try figuring this out. if you can't peek into chapters 3.1 and 3.2 and/or the slides]?
- First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results
- Then implement a quadratic approximation (hint check paragraph 2.4.2!).
- N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)

## Question 1 - Ricardo's knowledge
```{r}
# loading packages
library("rstan")

pacman::p_load(rethinking, ggplot2, dplyr, tidyverse, tidyr, gridExtra, tidybayes)

#set seed for consistency

set.seed(69)

### Riccardo's estimated knowledge

##grid aproximation##

# define grid
p_grid <- seq(from=0 , to=1 , length.out=1000)

## Priors
uni_prior <- dunif(p_grid, 0,1)
stan_prior <- dnorm(p_grid, mean = 0.8, sd = 0.2)

# compute likelihood at each value in grid
ric_likelihood <- dbinom(3 , size=6 , prob=p_grid )

# compute product of likelihood and prior
uns_ric_posterior <- ric_likelihood * uni_prior
ric_posterior <- uns_ric_posterior/sum(uns_ric_posterior)


# drawing samples
ric_samples <- sample( p_grid , prob=ric_posterior , size=1e4 , replace=TRUE )

#posterior
sum( ric_posterior[ p_grid > 0.5 ] ) # 50 % probability that Ricardo's knowledge above 0.5
p_grid[which.max(ric_posterior)]


## quadratic approximation
ric.qa <- rethinking::map(
  alist(
    w ~ dbinom(6,p) , # binomial likelihood
    p ~ dunif(0,1)), # uniform prior
    data=list(w=3) )

# display summary of quadratic approximation
precis( ric.qa )
#curve( dnorm( x , 0.5 , 0.2 ) , lty=2 , add=TRUE )
```

## Question 2
2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel.
2a. Produce plots of the prior, and posterior for each teacher.


```{r}

## posterior
calc_teacher_knowledge <- function(n_correct, n_question, prior, length_out = 10000){
  p_grid <- seq(0, 1, length.out= length(prior))
  
  likelihood <- dbinom(n_correct, size = n_question, prob=p_grid)
  
  unstd.posterior <- prior*likelihood
  
  posterior <- unstd.posterior/sum(unstd.posterior)
  
  sampl <- sample(p_grid, prob = posterior, size=1e4 , replace=TRUE)
  
  max_posterior <- p_grid[which.max(posterior)]
  
  HPDI <- HPDI(sampl,prob=0.8) # if bell-shaped, prob doesn't matter
  SD <- sd(sampl)
  
  return(list(teacher_posterior = posterior, 
              likelihood = likelihood,
              max_posterior = max_posterior, 
              teacher_HPDI = HPDI,
              sample = sampl,
              SD=SD))
}

## plotting

pretty_plot <- function(p_grid, prior, posterior, title = " "){
  # define data
  d <- tibble(p_grid = p_grid, 
              prior = prior/length(p_grid),
              posterior = posterior)

  # make to long format
  d <- d %>% 
    pivot_longer(cols = c("prior", "posterior"), names_to = "name", values_to = "value")
  
  # make a 
  p <- ggplot(d, aes(x = p_grid, y = value, color = name)) + 
    geom_line() + 
    labs(x = "CogSci Knowledge (p)", y = "Density", title = title) + 
    theme_bw() + 
    ggplot2::theme(panel.background = element_rect(fill = "White"),
                   panel.border = element_blank()) +
    scale_colour_brewer(palette = "Dark2", direction = 1)
  return(p)
}

#function for giving summary stats
summary_stats <- function(teacher){
  print(teacher$teacher_HPDI)
  print(teacher$max_posterior)
  print(teacher$SD)
}

## Riccardo
ric <- calc_teacher_knowledge(3, 6, uni_prior)
ric_plot <- pretty_plot(p_grid, uni_prior, ric$teacher_posterior, title = "Riccardo's estimated knowledge")
summary_stats(ric)

## Kristian
kris <- calc_teacher_knowledge (2, 2, uni_prior)
kris_plot <- pretty_plot(p_grid, uni_prior, kris$teacher_posterior, title = "Kristian's estimated knowledge" )
summary_stats(kris)

## Josh
josh<- calc_teacher_knowledge(160, 198, uni_prior)
josh_plot <- pretty_plot(p_grid, uni_prior, josh$teacher_posterior, title = "Josh's estimated knowledge" )
summary_stats(josh)

## Mikkel
mik <- calc_teacher_knowledge(66,132, uni_prior)
mik_plot <- pretty_plot(p_grid, uni_prior, mik$teacher_posterior, title = "Mikkel's estimated knowledge" )
summary_stats(mik)

## plotting knowledge of everyone at one
grid.arrange(ric_plot, kris_plot, josh_plot, mik_plot, nrow=2)

sum(kris$sample<josh$sample)/1e4*100 #josh is smarter than Kristian
```
## Question 3
3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?
3a. Produce plots of the prior and posterior for each teacher.
```{r}
## Riccardo
ric3 <- calc_teacher_knowledge (3, 6, stan_prior)
ric3_plot <- pretty_plot(p_grid, stan_prior, ric3$teacher_posterior, title = "Riccardo's estimated knowledge" )
summary_stats(ric3)


## Kristian
kris3 <- calc_teacher_knowledge (2, 2, stan_prior)
kris3_plot <- pretty_plot(p_grid, stan_prior, kris3$teacher_posterior, title = "Kristian's estimated knowledge" )
summary_stats(kris3)

## Josh
josh3 <- calc_teacher_knowledge(160, 198, stan_prior)
josh3_plot <- pretty_plot(p_grid, stan_prior, josh3$teacher_posterior, title = "Josh's estimated knowledge" )
summary_stats(josh3)

## Mikkel
mik3 <- calc_teacher_knowledge(66,132, stan_prior)
mik3_plot <- pretty_plot(p_grid, stan_prior, mik3$teacher_posterior, title = "Mikkel's estimated knowledge" )
summary_stats(mik3)

## plotting
grid.arrange(ric3_plot, kris3_plot,josh3_plot, mik3_plot, nrow=2)
```

## Question 4
4. You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results? Why?

```{r}
## Ricardo
ric4_uni <- calc_teacher_knowledge (300, 600, uni_prior)
ric4_stan <- calc_teacher_knowledge(300, 600, stan_prior)


## Kristian
kris4_uni <- calc_teacher_knowledge (200, 200, uni_prior)
kris4_stan <- calc_teacher_knowledge(200, 200, stan_prior)


## Josh
josh4_uni <- calc_teacher_knowledge (16000, 19800, uni_prior)
josh4_stan <- calc_teacher_knowledge(16000, 19800, stan_prior)


## Mikkel
mik4_uni <- calc_teacher_knowledge (6600, 13200, uni_prior)
mik4_stan <- calc_teacher_knowledge(6600, 13200, stan_prior)

#summary stats with uniform priors
summary_stats(ric4_uni)
summary_stats(kris4_uni)
summary_stats(josh4_uni)
summary_stats(mik4_uni)

#summary stats with an informed prior
summary_stats(ric4_stan)
summary_stats(kris4_stan)
summary_stats(josh4_stan)
summary_stats(mik4_stan)


## Riccardo
ric4_stan_plot <-
  pretty_plot(p_grid, stan_prior, ric4_stan$teacher_posterior, title = "Riccardo's estimated knowledge")


## Kristian

kris4_stan_plot <-
  pretty_plot(p_grid, stan_prior, kris4_stan$teacher_posterior, title = "Kristian's estimated knowledge")

## Josh

josh4_stan_plot <-
  pretty_plot(p_grid, stan_prior, josh4_stan$teacher_posterior, title = "Josh's estimated knowledge")

## Mikkel

mik4_stan_plot <-
  pretty_plot(p_grid, stan_prior, mik4_stan$teacher_posterior, title = "Mikkel's estimated knowledge")

## plotting
grid.arrange(ric4_stan_plot, kris4_stan_plot,josh4_stan_plot, mik4_stan_plot, nrow=2, top = "with 100 times more data and informed priors")

#Riccardo
ric4_uni_plot <-
  pretty_plot(p_grid, stan_prior, ric4_uni$teacher_posterior, title = "Riccardo's estimated knowledge")


## Kristian

kris4_uni_plot <-
  pretty_plot(p_grid, stan_prior, kris4_uni$teacher_posterior, title = "Kristian's estimated knowledge")

## Josh

josh4_uni_plot <-
  pretty_plot(p_grid, stan_prior, josh4_uni$teacher_posterior, title = "Josh's estimated knowledge")

## Mikkel

mik4_uni_plot <-
  pretty_plot(p_grid, stan_prior, mik4_uni$teacher_posterior, title = "Mikkel's estimated knowledge")

## plotting
## plotting
grid.arrange(ric4_uni_plot, kris4_uni_plot,josh4_uni_plot, mik4_uni_plot, nrow = 2, top = "With 100 times more data and uniform priors") 


```

## Question 5
5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?


### Second part: Focusing on predictions


Questions to be answered (but see guidance below):
1- Write a paragraph discussing how assessment of prediction performance is different in Bayesian vs. frequentist models
2- Provide at least one plot and one written line discussing prediction errors for each of the teachers.

This is the old data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

This is the new data:
- Riccardo: 9 correct answers out of 10 questions (then he freaks out about teaching preparation and leaves)
- Kristian: 8 correct answers out of 12 questions
- Josh: 148 correct answers out of 172 questions (again, Josh never gets bored)
- Mikkel: 34 correct answers out of 65 questions


```{r}
## calculating new knowledge
ric_new <- calc_teacher_knowledge(9, 10, ric3$teacher_posterior)
kris_new <- calc_teacher_knowledge(8, 12, kris3$teacher_posterior)
josh_new <- calc_teacher_knowledge(148, 172, josh3$teacher_posterior)
mik_new <- calc_teacher_knowledge(34, 65, mik3$teacher_posterior)


## function for a plot of posterior comparison
comp_distrib_plot <- function(old_posterior, new_posterior, title = ""){
  # define data
  d <- tibble(old_posterior = old_posterior,
              new_posterior = new_posterior,
              p_grid = p_grid)
  d <- d %>% 
      pivot_longer(cols = c("old_posterior", "new_posterior"), names_to = "name", values_to = "value")
  
  # make a 
  p <- ggplot(d, aes(x = p_grid, y = value, colour = name)) + 
    geom_line()+ 
    labs(y = "Density", x = " Posterior of teacher's knowledge", title = title)
  p
  
  return(p)
}

## running the plot comparison function on data of all 4 teachers
ric_comp <- comp_distrib_plot(ric3$teacher_posterior, ric_new$teacher_posterior, title = "Riccardo's estimated knowledge 2019 & 2020")
kris_comp <- comp_distrib_plot(kris3$teacher_posterior, kris_new$teacher_posterior, title = "Kristian's estimated knowledge 2019 & 2020")
josh_comp <- comp_distrib_plot(josh3$teacher_posterior, josh_new$teacher_posterior, title = "Josh's estimated knowledge 2019 & 2020")
mik_comp <- comp_distrib_plot(mik3$teacher_posterior, mik_new$teacher_posterior, title = "Mikkel's estimated knowledge 2019 & 2020")

all_comparison <- grid.arrange(ric_comp, kris_comp, josh_comp, mik_comp, nrow=2)
 
### histogram function

hist_function <- function(posterior, n_questions, bin_width, x = "", title = ""){
  
  #drawing samples
  sampling <- sample(size = 100000,
                     x = p_grid,
                     prob = posterior,
                     replace = T)
  pos_pred <- rbinom(100000,
                     size = n_questions,
                     prob = sampling)
  #doing some nice plots
  d <- as.data.frame(pos_pred)
 
   g <- ggplot(d, aes(pos_pred, fill = "pink"))+ 
     geom_histogram() + 
     labs(x = x, y = "Count", title = title) +
     stat_bin(binwidth = bin_width)
   return(g)
}

ric_his <- hist_function(ric3$teacher_posterior, 10, 1, x = "Correct answers (out of 10)", title = "Predictions of Riccardo")
kris_his <- hist_function(ric3$teacher_posterior, 12, 1, x = "Correct answers (out of 12)", title = "Predictions of Kristian")
josh_his <- hist_function(ric3$teacher_posterior, 172, 1, x = "Correct answers (out of 172)", title = "Predictions of Josh")
mik_his <- hist_function(ric3$teacher_posterior, 65, 1, x = "Correct answers (out of 65)", title = "Predictions of Mikkel")

grid.arrange(ric_his, kris_his, josh_his, mik_his, nrow=2)


# this function makes a histogram of errors
error_hist <- function(posterior, questions, correct, name){
  samples <-  sample(p_grid , prob=posterior , size=100000 , replace=TRUE )
  error_hist <- hist(samples * questions - correct,  main=name, breaks=questions, xlab = "Error Count")
  return(error_hist)
}

# this function makes a histogram of errors (used for large samples to control bin width)
error_hist2 <- function(posterior, questions, correct, name){
  samples <-  sample(p_grid , prob=posterior , size=100000 , replace=TRUE)
  error_hist <- hist(samples * questions - correct,  main=name, xlab = "Error Count")
  return(error_hist)
}
error_hist(ric3$teacher_posterior,10,9, "Riccardo Error")
error_hist(kris3$teacher_posterior,12,8, "Kristian Error")
error_hist2(josh3$teacher_posterior,172,148, "Josh Error")
error_hist2(mik$teacher_posterior,65,34, "Mikkel Error")


